Community-Conditioned Language Models

### 2021-06-04 paper/paper.tex
improve presentation of embedding correlatino

Computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
solution is
\[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). That is, each
singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).

@@

The \(d(L,S)\) metric ranges from \(0\) 
(corresponding to perfect correlation, 
obtained for example if \(L=S\)) 
to \(n\) 
(corresponding to perfect independence), 
where $n=510$ is the number of communities considered.

To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
We have experimentally computed \(d(L,S)\) for five random  ``embeddings'' 
with the same shape as the CCLM embeddings.
We observed a mean of \(μ_d=431.39\) and a (Bessel's-corrected) 
standard deviation \(s_d=2.90\) in their distance from the social embedding, \(S\). %\jp{Unfortunate overloading} luckily it seems that sample stdev is often written with an $s$ to differentiate from population stdev

Thus if \(d(L,S)\) is below the mean by several standard deviations, we can safely
assume that there is statistically significant correlation between
\(L\) and \(S\).  This is the case: we observe a difference between 61 and 68 standard deviations
(\cref{tab:embedding-correlations}). Further, it is clear that the CCLM embedding
capture some of the same information as the social-network embeddings.
We also give a sense of \emph{how} the correlation is manifested by
analysis of the two principal components of correlation in the
linguistic embeddings, \(U_0\) and \(U_1\). To do so we plot the
projection of embeddings on along \(U_0\) and \(U_1\), together with
the corresponding singular values.\jp{TODO}

>>>>

Computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
solution is
\[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
decomposition \(U^TΣV = LS^T\). This way, the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). That is, each
singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).
Additionally, assuming a test community \(c\) and a CCLM embedding $x_c$ for it, we can make a prediction
of its social embedding to be $\hat y_c = V^TUx_c$ and expect a cosine similarity with the actual social embedding $y_c$ to be on average
$1 - d(L,S) / n$.

@@

The \(d(L,S)\) metric ranges from \(0\) (corresponding to perfect
correlation, obtained for example if \(L=S\)) to \(n\) (corresponding
to perfect independence), where $n=510$ is the number of communities
considered.
%
For our various embiddings, we observe a distance \(d(L,S)\) between
232 and 354, which corresponds to an expected cosine similarity beween
0.5 and 0.54. This strongly suggests that much information is shared
between social-network embeddings and linguistic embeddings.

However, to get a more precise sense of whether this correlation is
statistically significant, we have experimentally computed \(d(L,S)\)
for 10 random ``embeddings'' with the same shape as the CCLM
embeddings.  We observed a mean of \(μ_d=431.39\) and a
(Bessel's-corrected) standard deviation \(s_d=2.90\) in their distance
from the social embedding,
\(S\). %sigma vs. s notation: it seems that sample stdev is often written with an $s$ to differentiate from population stdev

Because the value of \(d(L,S)\) that we get for actual embedding is
between 61 and 68 standard deviations
(\cref{tab:embedding-correlations}), we can safely conclude that there
is statistically significant correlation between \(L\) and \(S\).

% Further, it is clear that the CCLM embedding
% capture some of the same information as the social-network embeddings.
% We also give a sense of \emph{how} the correlation is manifested by
% analysis of the two principal components of correlation in the
% linguistic embeddings, \(U_0\) and \(U_1\). To do so we plot the
% projection of embeddings on along \(U_0\) and \(U_1\), together with
% the corresponding singular values.
